### 2023-06-15 manuscript/chapters/3_mappings.qmd
pwoofwead !!!1

In the previous chapter, we have established a simple toy model for a neural field in $p$ dimensions, and have shown how it can be embedded in $[0,1]^p$. In this chapter, we discuss in-depth what types of mappings can be used to map the $p$-dimensional space to $[0,1]$, and in particular first study the case $p=2$. Before getting into too much detail, let us give some insight as to why this problem is not trivial.

@@

Let us for a moment consider another type of model for networks of neurons, that of homogeneous population dynamics, that describes how $N_\text{pop}$ different populations interact with each other. These types of models emerge when networks of neurons exhibit a finite number of "neuron types". We then group similar neurons ad make the approximation that all neurons inside the population have the same potential, and the same interaction weights with other neurons. In effect, every population then represents in actuality an infinity of discrete identical neurons.

The result of this is that the population dynamics can be written with the same equation as a finite network of neurons, but it already represents the $N \to \infty$ limit. We write the potential of the populations $H_a(t)$ and the population connectivity $\tilde J_{ab}$, where $a,b \in \{1, \cdots, N_\text{pop}\}^2$ in @eq-pop-dynamics.

@@

We don't show the derivation of population dynamics here (see @GerKis14, chapter 12.3 for a detailed analysis), but we point out the implicit scaling behavior of $\tilde J_{ab}$. Letting $i$ be a neuron of population $a$ and $j$ be a neuron of population $b$, the connectivity between the two neurons is $J_{ij}$ and the population connectivity scales with the number of neurons in the presynaptic population $|b|$. Furthermore, we make the reasonable assumption that $J_{ij}$ scales as $J^0_{ij} / N$, where $J^0_{ij} \sim \bO(1)$.

$$
\tilde J_{ab} = \frac{|b|}{N} J^0_{ij}
$$

@@

When taken in isolation, the previous developments might sound a little bit silly, though they serve to give an impression of the problem at hand, and an interpretation of the rest of this work. For a finite number of populations $N_\text{pop}$, the population dynamics is perfectly defined, however what happens as $N_\text{pop} \to \infty$ ? The core of the question lies in the fact that we can easily *enumerate* a finite number of populations, but the sum $\sum_{b=1}^{N_\text{pop}}$ must have a clearly defined interpretation when this number tends to infinity.

@@

Additionally, the notion of population dynamics helps give another interpretation of the grid integration method given in @sec-gridmethod. When we discretize the $[0,1]^2$ space, we are effectively performing a coarse-graining approximation and simulating a finite number of populations. For every bin localized by $(v_1, v_2)$, we are sampling the connectivity kernel $w_U((v_1, v_2), \cdot)$ and 2D neural field $h_U((v_1, v_2), t)$ (see @fig-grid-sampling). In the rest of the chapter, we refer to these 2D bins as "square populations", and to the corresponding 1D bins as "segment populations".

![Every point of the grid effectively performs coarse-graining, and defines a point where we sample the connectivity kernel $w_U$ and the 2D neural field $h_U$](figures/photo_2023-06-10_21-13-23.jpg){#fig-grid-sampling}

@@

Therefore, we can see why the composition of $w_U$ with a possibly non-differentiable mapping $S$ might be problematic. Since the only thing we know about $S$ is that it is measurable and bijective, *there is no guarantee that the regularity of the kernel is conserved through the composition with the mapping $S$.*

@@

Perhaps the linear projection or CDF examples are a bit too naive, and we need to look for more complex examples. We might ask the question of whether a diffeomorphism (a differentiable bijection with a differentiable inverse) even exists between the unit square and the unit segment. Unfortunately, results from topology give a negative answer, and even worse, it can be shown that there is no homeomorphism (a continuous bijection with continuous inverse) between the unit square and the unit segment. This is stated by Netto's theorem, which imposes continuous bijections to conserve the dimensionality between the domain and the image.

@@

1. Cantor functions are bijections between $[0,1]$ and $[0,1]^2$. By the non-existence theorem, these functions cannot be continuous.
2. Peano functions are continuous surjections from $[0,1]$ to $[0,1]^2$. They cannot be injections. For reasons that will become apparent in the following sections, Peano functions are often called space-filling curves.

In light of these results, we might be tempted to search for expressions for Cantor-type functions, since having a bijection would allow us to have a direct mapping between the square and the segment populations. Unfortunately, these functions are discontinuous which breaks the regularity of the kernel.

Peano functions seem like good candidates for the mapping $S^{-1}$, and are extensively studied in the literature (we refer for instance to [@sagan_space-filling_1994;@Sagan1993]). They have the nice property that they are continuous and surjective, so we the segment populations can "reach" the entirety of the square populations. All things considered, non-injectivity is not a problem, since it just means that two segment populations can "refer" to the same square population, which does not break the regularity of the kernel, since the two segment populations can just have their own neighborhoods. Furthermore, we will see that in our examples, the non-injectivity can be ignored because the mapping is quasi-bijective.

@@

We consider $(v_1, v_2) \in [0,1]^2$. Given $n \in \mathbb N$, we write a (truncated) binary expansion of $v_1$ and $v_2$.

@@

The intuition behind this expansion is we take a segment, split it in two, write 0 to go left, write one to go right, then recurse. Then combine the bits of $v^{(n)}_1$ and $v^{(n)}_2$ to get the position in the $[0,1]^2$ embedding.

![Splitting $[0,1]$ recursively gives the binary bit expansion of $v_1$](figures/photo_2023-06-10_17-21-28.jpg)

@@

In general, each bit $b^k$ of the image would be a binary function of the $n$ bits of $v_1$ and the $n$ bits of $v_2$, implicitly $b_k = b_k(b^1_1, b^1_2, \cdots, b^1_n, b^2_1, b^2_2, \cdots, b^2_n)$.
With this, we can formulate any mapping of $4^n$ distinct elements (here : squares of size $2^{-n} \times 2^{-n}$) onto $4^n$ elements (here : segments of length $4^{-n}$).

The intuition behind the use of $2n$ bits for the image is that we take $2^n \times 2^n = 4^n$ squares in $[0,1]^2$ and map those to $4^n$ segments in $[0,1]$. Informally, we map $n+n$ bits onto $2n$ bits.

In the following, we will restrict ourselves to mappings where the functions $b_k(\{b^1_l\}_{l\in\{1,\cdots,n\}}, \{b^2_l\}_{l\in\{1,\cdots,n\}})$ are defined in such a way that each bit of the inputs is used once, and only once. Therefore, for each bit $k$ there is a function $b_k(b^{\mu_k}_{l_k})$ acting on only one bit of the input. $\mu_k \in \{1,2\}$ is the input component and $l_k \in \{1,\cdots,n\}$ the bit number corresponding to $k$. Since this function is binary, the only allowed definitions are

@@

For finite $n$, we write the inverse of the column mapping in @eq-column-inverse.

$$
\begin{aligned}
v^{(n)}_1 = C^{-1}_1(\alpha^{(n)}) &= 0.b_1 b_2 \cdots b_n = \sum_{k=1}^n b_{k} 2^{-k} \\
v^{(n)}_2 = C^{-1}_2(\alpha^{(n)}) &= 0.b_{n+1} b_{n+2} \cdots b_{2n} = \sum_{k=1}^n b_{n+k} 2^{-k}
\end{aligned}
$$ {#eq-column-inverse}

This way of enumerating the square populations might be familiar to people who have worked with numerical arrays. In such (contiguous) arrays, the values are laid out in memory in "row-major" order (C-style) or "column-major" order (Fortran-style); and we can go from the 1D memory representation to a 2D matrix representation by using "reshape" operations. Effectively, this is how this mapping is implemented in code, using `numpy.ravel_multi_index` (2D to 1D) and `numpy.unravel_index` (1D to 2D).

@@

Research in computer science and efficient data structures has given rise to other ways of storing 2-dimensional information in a 1-dimensional memory.
A problem in GPU computing is that of texture locality: sampling textures often involves reading data corresponding to a small 2D region, and therefore to improve speed (minimize cache misses), the 2D data should be packed close together in the 1D memory.
One way of ordering the 2D texture data in the 1D memory is by using the Z-order curve (also known as "Morton mapping" or "Lebesgue curve" @Sagan1994). In this way, points close in 2D *tend* to be tightly packed in 1D. We can define the Z-order curve (Z-mapping) in the following way[^zcurve] :

@@

And its inverse is :

$$
\begin{aligned}
v^{(n)}_1 = Z^{-1}_1(\alpha^{(n)}) &= 0.b_1 b_3 \cdots b_{2n-1} = \sum_{k=1}^n b_{2k-1} 2^{-k} \\
v^{(n)}_2 = Z^{-1}_2(\alpha^{(n)}) &= 0.b_2 b_4 \cdots b_{2n} = \sum_{k=1}^n b_{2k} 2^{-k}
\end{aligned}
$$ {#eq-z-inverse}

@fig-z is a visualization for the Z-mapping when $n=2$, and shows how it is composed of the "Z" shapes giving it its name.

@@

```{python}
#| label: fig-anti-z
#| fig-cap: Anti-Z mapping for $n=2$. Numbers represent the ordering of the populations in the $[0,1]$ embedding. We don't plot the trace, because it has many crossovers and overlaps which makes the plot hard to read.
#| echo: false
fig, ax = plot_trace(AntiZMapping, nrec=2, do_trace=False)
display(fig)
```

@fig-anti-z gives some insight into the problem of this mapping. For a fixed input $(v_1,v_2)$, we see that the output point position jumps around and that these jumps do not decay, they are always of order 1. The intuition behind pointwise convergence is that as $n$ increases, the image of a point by $S^n$ is continually refined.

Another way of looking at this is to consider how $\alpha^{(n)}$ fluctuates as the precision of the input increases with $n \to \infty$. As we refine the input with fluctuations of order $2^{-n}$, the output $\alpha^{(n)} = b^1_n 4^{-1} + b^2_n 4^{-2} + b^1_{n-1} 4^{-3} + b^2_{n-1} 4^{-4} \cdots$ fluctuates (at least) with amplitude $\sim 4^{-2}$ independantly of $n$. Therefore the output never converges, and the mapping $A^{n}$ is not pointwise convergent.

@@

The limit is not bijective, because the bits of $v_2$ are "lost" in the $n\to\infty$ limit, and the inverse $C_2^{-1}(\alpha)$ is not well-defined, since it "takes bits from $\infty$ to $2\infty$".

The Z-mapping however seems to converge to a fractal curve. Numerically, there seems to be an infinite number of discontinuities, but the size of these discontinuities is overwhelmingly small, and *on average* the size of these numerical discontinuities vanishes as $n\to\infty$. We won't go further into detail about this in this section (in @sec-zcurve-lebesgue we develop the convergence to a space-filling fractal curve), but this gives some intuition for the arguments presented in @sec-locality.

@@

These two conditions do not leave a lot of flexibility in the design of new mappings. We can evidently define mappings where we interleave the bits in a different order, such as $0.b^2_1 b^1_1 b^2_2 b^1_2 \cdots$; or mappings where we reorder the $k$ bits of the input, as in $0.b^1_2 b^2_2 b^1_1 b^2_1 b^1_4 b^2_4 b^1_3 b^2_3 \cdots$, but the conditions restrict the reordering to be permutations of neighbouring bits. In essence, these mappings are not fundamentally different, and the Z-mapping can be seen as a "canonical representation" of the pointwise convergent mappings with bijective limits.

@@

## Coarse-graining the neural field in $[0,1]$

@@

In the previous sections, we have discussed what mappings have good properties, and have given a few hints as to how we can implement them numerically. Since the mappings we presented can be described as a binary expansion, they naturally lend themselves to numerical approximations. Furthermore, this is made especially easy since the mappings we defined simply are permutations of bits (in the $n < \infty$ case), and so we can write a one-to-one mapping between (unsigned integers).

@@

In the continuation of the simulation methods established in the @sec-simnf, we adopt a grid discretization scheme, as it naturally lends itself to be applied with the mappings. We however note that using binning methods, the Monte-Carlo simulation can still be applied. The intuition behind this is that sampling populations inside a small square of size $2^{-n} \times 2^{-n}$ gives rise to the notion of an "average connectivity kernel", and in our case the notion of "mean patterns". More details are given in @sec-mean-patterns.

@@

### Iterative application of $S$ : from $[0,1]^3$ to $[0,1]$

>>>>

In the previous chapter, we have established a simple toy model for a neural field in $p$ dimensions, and have shown how it can be embedded in $[0,1]^p$. In this chapter, we discuss in-depth what types of mappings can be used to map the $p$-dimensional space to $[0,1]$, and in particular we study the case $p=2$. Before getting into too much detail, let us give some insight as to why this problem is not trivial.

@@

Let us for a moment consider another type of model for networks of neurons, that of homogeneous population dynamics, that describes how $N_\text{pop}$ different populations interact with each other. These types of models emerge when networks of neurons exhibit a finite number of "neuron types". We then group neurons of the same type, and make the approximation that they have identical potentials and the same interaction weights with other neurons. In effect, every population then represents an infinite number of discrete identical neurons.

This results in writing population dynamics with the same equation as a finite network of neurons. However, the equation *already represents* the $N \to \infty$ limit. We write the potential of the populations $H_a(t)$ and the population connectivity $\tilde J_{ab}$, where $a,b \in \{1, \cdots, N_\text{pop}\}^2$ in @eq-pop-dynamics.

@@

We don't show the derivation of population dynamics here (see @GerKis14, chapter 12.3 for a detailed analysis), but we point out the implicit scaling behavior of $\tilde J_{ab}$. Letting $i$ be a neuron of population $a$ and $j$ be a neuron of population $b$, the connectivity between the two neurons is $J_{ij}$, and the population connectivity scales with the number of neurons in the presynaptic population $|b|$. Furthermore, we make the reasonable assumption that $J_{ij}$ scales as $J^0_{ij} / N$, where $J^0_{ij} \sim \bO(1)$. We then have

$$
\tilde J_{ab} = \frac{|b|}{N} J^0_{ij}.
$$

@@

When taken in isolation, the previous developments might sound a little bit silly, though they serve to give an impression of the problem at hand, and an interpretation for the rest of this work. For a finite number of populations $N_\text{pop}$, the population dynamics is well-defined, however, what happens as $N_\text{pop} \to \infty$ ? The core of the question lies in the fact that we can easily *enumerate* a finite number of populations, but the sum $\sum_{b=1}^{N_\text{pop}}$ must have a clearly defined interpretation when this number tends to infinity.

@@

Additionally, the notion of population dynamics helps give another interpretation of the grid integration method given in @sec-gridmethod. When we discretize the $[0,1]^2$ space, we are effectively performing a coarse-graining approximation and simulating a finite number of populations. For every bin localized by $(v_1, v_2)$, we are sampling the connectivity kernel $w_U(\cdot, (v_1, v_2))$ and 2D neural field $h_U((v_1, v_2), t)$ (see @fig-grid-sampling). In the rest of the chapter, we refer to these 2D bins as "square populations", and to the corresponding 1D bins obtained by the mapping $S$ as "segment populations".

![(TODO, reorder args of $w_U$) Every point of the grid effectively performs coarse-graining, and defines a point where we sample the connectivity kernel $w_U$ and the 2D neural field $h_U$](figures/photo_2023-06-10_21-13-23.jpg){#fig-grid-sampling}

@@

For these reasons, we can see why the composition of $w_U$ with a possibly non-differentiable mapping $S$ might be problematic. Since the only thing we know about $S$ is that it is measurable and bijective, *there is no guarantee that the regularity of the kernel is conserved through composition with the mapping $S$.*

@@

Perhaps these examples are a bit too naive, and we need to look for more complex mappings. We might ask the question of whether a diffeomorphism (a differentiable bijection with a differentiable inverse) even exists between the unit square and the unit segment. Unfortunately, results from topology give a negative answer, and even worse, it can be shown that there is no homeomorphism (a continuous bijection with continuous inverse) between the unit square and the unit segment. This is stated by Netto's theorem, which imposes continuous bijections to conserve the dimensionality between the domain and the image.

@@

1. Cantor functions are bijections between $[0,1]$ and $[0,1]^2$. By Netto's theorem, these functions cannot be continuous.
2. Peano functions are continuous surjections from $[0,1]$ to $[0,1]^2$. They cannot be injections. For reasons that will become apparent in the following sections, Peano functions are often called space-filling curves.

In light of these results, we might be tempted to search for expressions of Cantor-type functions, since having a bijection would allow us to have a direct mapping between the square and the segment populations. Unfortunately, these functions are discontinuous which breaks the regularity of the kernel.

Peano functions seem like good candidates for the mapping $S^{-1}$, and are extensively studied in the literature around space-filling curves (we refer for instance to [@sagan_space-filling_1994]). They have the nice property that they are continuous and surjective, so we the segment populations can "reach" the entirety of the square populations. All things considered, non-injectivity is not a problem, since it just means that two segment populations can "refer" to the same square population, which does not break the regularity of the kernel, since the two segment populations can just have their own neighborhoods. Furthermore, we will see that in our examples, the non-injectivity can be ignored because the mapping is "quasi-bijective".

@@

Previously, we motivated $S$ as the limit of functions $S^n$ acting on finer and finer grids. Let us consider $(v_1, v_2) \in [0,1]^2$. Given $n \in \mathbb N$, we write a (truncated) binary expansion of $v_1$ and $v_2$.

@@

The intuition and illustration @fig-recurse-binary behind this expansion is that we locate each component by recursively splitting the segment in two sub-segments of equal length. At every step, taking the left or right segments defines the sequence of binary digits.
<!-- Then combine the bits of $v^{(n)}_1$ and $v^{(n)}_2$ to get the position in the $[0,1]^2$ embedding. -->

![(TODO rename bins to segments) Splitting $[0,1]$ recursively gives the binary bit expansion of $v_1$](figures/photo_2023-06-10_17-21-28.jpg){#fig-recurse-binary}

@@

In general, each bit $b^k$ of the image would be a binary function of the $n$ bits of $v_1$ and the $n$ bits of $v_2$, implicitly $b_k = b_k(b^1_1, b^1_2, \cdots, b^1_n, b^2_1, b^2_2, \cdots, b^2_n)$.
With this, we can formulate any mapping between sets of $4^n$ distinct elements.

The intuition behind the use of $2n$ bits for the image is that we take $2^n \times 2^n = 4^n$ squares of size $2^{-n} \times 2^{-n}$ in $[0,1]^2$, and map those to $4^n$ segments of size $4^{-n}$ in $[0,1]$. Informally, we map $n+n$ bits onto $2n$ bits.

In the following, we will restrict ourselves to mappings where the functions $b_k$ are defined in such a way that each bit of the inputs $v_1$ and $v_2$ is used only once. Therefore, for each bit $k$ of the output, there is a function $b_k(b^{\mu_k}_{l_k})$ acting on only one bit of the input. $\mu_k \in \{1,2\}$ is the input component and $l_k \in \{1,\cdots,n\}$ the bit number corresponding to $k$. Since this function is binary, the only allowed definitions are

@@

For finite $n$, we write the inverse of the column mapping:

$$
\begin{aligned}
v^{(n)}_1 = C^{-1}_1(\alpha^{(n)}) &= 0.b_1 b_2 \cdots b_n = \sum_{k=1}^n b_{k} 2^{-k} \\
v^{(n)}_2 = C^{-1}_2(\alpha^{(n)}) &= 0.b_{n+1} b_{n+2} \cdots b_{2n} = \sum_{k=1}^n b_{n+k} 2^{-k}.
\end{aligned}
$$ {#eq-column-inverse}

This way of enumerating the square populations might be familiar to people who have worked with numerical arrays. In such (contiguous) arrays, the values are laid out in memory in "row-major" order (C-style) or "column-major" order (Fortran-style); and we can go from the 1D memory representation to a 2D matrix representation by using "reshape" operations. Effectively, this is how this mapping is implemented in code, using [`numpy.ravel_multi_index`](https://numpy.org/doc/stable/reference/generated/numpy.ravel_multi_index.html) (2D to 1D) and [`numpy.unravel_index`](https://numpy.org/doc/stable/reference/generated/numpy.unravel_index.html) (1D to 2D).

@@

Research in computer science and efficient data structures has given rise to other ways of storing 2-dimensional information in a 1-dimensional memory.
A problem in GPU computing is that of texture locality: sampling textures often involves reading data corresponding to a small 2D region, and therefore to improve speed (minimize cache misses), the 2D data should be packed close together in the 1D memory.
One way of ordering the 2D texture data in the 1D memory is by using the Z-order curve (also known as "Morton mapping" or "Lebesgue curve" @Sagan1994). In this way, points close in 1D *tend* to be tightly packed in 2D. We can define the Z-order curve (Z-mapping) in the following way:[^zcurve]

@@

And its inverse is:

$$
\begin{aligned}
v^{(n)}_1 = Z^{-1}_1(\alpha^{(n)}) &= 0.b_1 b_3 \cdots b_{2n-1} = \sum_{k=1}^n b_{2k-1} 2^{-k} \\
v^{(n)}_2 = Z^{-1}_2(\alpha^{(n)}) &= 0.b_2 b_4 \cdots b_{2n} = \sum_{k=1}^n b_{2k} 2^{-k}.
\end{aligned}
$$ {#eq-z-inverse}

@fig-z is a visualization of the Z-mapping when $n=2$, and shows how it is composed of the "Z" shapes giving it its name.

@@

```{python}
#| label: fig-anti-z
#| fig-cap: Anti-Z mapping for $n=2,3,4$. Numbers represent the ordering of the populations in the $[0,1]$ embedding. We don't plot the trace, because it has many crossovers and overlaps which makes the plot hard to read.
#| fig-subcap: 
#|   - $n=2$
#|   - $n=3$
#|   - $n=4$
#| layout-ncol: 3
#| echo: false
fig, ax = plot_trace(AntiZMapping, nrec=2, do_trace=False)
display(fig)
fig, ax = plot_trace(AntiZMapping, nrec=3, do_trace=False)
display(fig)
fig, ax = plot_trace(AntiZMapping, nrec=4, do_trace=False, do_numbers=False)
display(fig)
```

@fig-anti-z gives some insight into the problem of this mapping. For a fixed input $(v_1,v_2)$, we see that the output point position jumps around when $n$ is changed, and that these jumps seem to always be of order 1. The intuition behind pointwise convergence is that as $n$ increases, the image of a point by $S^n$ is continually refined.

Consider how $\alpha^{(n)}$ fluctuates as the precision of the input increases with $n \to \infty$. As we refine the input with fluctuations of order $2^{-n}$, the output $\alpha^{(n)} = b^1_n 4^{-1} + b^2_n 4^{-2} + b^1_{n-1} 4^{-3} + b^2_{n-1} 4^{-4} \cdots$ fluctuates (at least) with amplitude $\sim 4^{-2}$ independantly of $n$. Therefore the output never converges, and the mapping $A^{n}$ is not pointwise convergent.

@@

The limit is not bijective, because the bits of $v_2$ are "lost" in the $n\to\infty$ limit, and the inverse $C_2^{-1}(\alpha)$ is not well-defined, since it "takes bits between $\infty$ and $2\infty$".

The Z-mapping seems to converge to a fractal curve. Numerically, there seems to be an infinite number of discontinuities, but the size of these discontinuities is overwhelmingly small, so that *on average* they vanish as $n\to\infty$. We won't go further into detail about this in this section (in @sec-zcurve-lebesgue we develop the convergence to a space-filling fractal curve), but this gives some intuition for the arguments presented in @sec-locality.

@@

These two conditions do not leave a lot of flexibility in the design of new mappings. We can evidently define mappings where we interleave the bits in a different order, such as $0.b^2_1 b^1_1 b^2_2 b^1_2 \cdots$; or mappings where we reorder the $k$ bits of the input, as in $0.b^1_2 b^2_2 b^1_1 b^2_1 b^1_4 b^2_4 b^1_3 b^2_3 \cdots$, but the conditions restrict the reordering to be permutations of neighbouring bits. In essence, these mappings are not fundamentally different, and the Z-mapping can be seen as a "canonical representation" of the pointwise convergent mappings with bijective limits, which can be expressed as the permutation of binary digits.

@@

## Coarse-graining the neural field in $[0,1]$ {#sec-coarse-graining}

@@

In the previous sections, we have discussed what mappings have good properties, and have given a few hints as to how we can implement them numerically. Since the mappings we presented can be written as a binary expansion, they naturally lend themselves to numerical approximations. Furthermore, this is made especially easy since the mappings we defined simply are permutations of bits (in the $n < \infty$ case), and so we can write a one-to-one correspondence between unsigned integers.

@@

In the continuation of the simulation methods established in the @sec-simnf, we adopt a grid discretization scheme, as it naturally lends itself to be applied with the mappings. We however note that using binning methods, the Monte-Carlo simulation can still be applied. The intuition behind this is that sampled populations inside a small square of size $2^{-n} \times 2^{-n}$ give rise to the notion of an "average connectivity kernel", and in our case the notion of "mean patterns". More details are given in @sec-mean-patterns.

@@

### Iterative application of $S$ : from $[0,1]^3$ to $[0,1]$ {#sec-simulations-3d-2d-1d}
